\documentclass[a4paper,10pt,3p,final,pdftex]{elsarticle}

\biboptions{sort&compress}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{algorithm,algorithmic}
\usepackage{url}
\usepackage{color}
\usepackage{tikz}
\usepackage{multirow}
%\usepackage{cite}

\newdefinition{remark}{Remark}
\newcommand {\RM}[1]{\mathrm{#1}}
\renewcommand{\Re}{\mathbb{R}}

\newcommand{\opal}{\textsc{OPAL}}
\newcommand{\ippl}{\textsc{IP$^2$L}}
%%\newcommand{\oursolver}{\textsc{SAAMG-PCG}}

\journal{Journal of Computational Physics}

\begin{document}

\begin{frontmatter}

\title{A Fast Parallel Poisson Solver on Irregular Domains Applied to Beam
  Dynamics Simulations}

\author[psi]{A.~Adelmann\corref{cor}}
\ead{andreas.adelmann@psi.ch}

\author[eth]{P.~Arbenz}
\ead{arbenz@inf.ethz.ch}

\author[psi,eth]{Y.~Ineichen}
\ead{ineichen@inf.ethz.ch}

\address[psi]{Paul Scherrer Institut, CH-5234 Villigen, Switzerland}
\address[eth]{ETH Z\"urich, Chair of Computational Science,
  Universit\"atsstrasse 6, CH-8092 Z\"urich, Switzerland}

\cortext[cor]{Corresponding author}



\begin{abstract}
  We discuss the scalable parallel solution of the Poisson equation
  within a Particle-In-Cell (PIC) code for the simulation of electron
  beams in particle accelerators of irregular shape.  The problem is
  discretized by Finite Differences.  Depending on the treatment of the
  Dirichlet boundary the resulting system of equations is symmetric or
  `mildly' nonsymmetric positive definite.  In all cases, the system is
  solved by the preconditioned conjugate gradient algorithm with
  smoothed aggregation (SA) based algebraic multigrid (AMG)
  preconditioning.  We investigate variants of the implementation of
  SA-AMG that lead to considerable improvements in the execution times.
  We demonstrate good scalability of the solver on distributed memory
  parallel processor with up to 2048 processors.  We also compare our
  iterative solver with an FFT-based solver that is more commonly used
  for applications in beam dynamics.
\end{abstract}

%MAX: 6 keywords
\begin{keyword}
  Poisson equation \sep Irregular domains \sep Preconditioned conjugate
  gradient algorithm \sep Algebraic multigrid \sep Beam dynamics \sep
  Space-charge
\end{keyword}

\end{frontmatter}

\nocite{*}
 
\input{intro}
\input{discretization}
\input{method}
\input{implementation}
\input{results}
\input{conclusion}

\section*{Acknowledgments}

The majority of computations have been performed on the Cray XT4 in the
framework of the PSI CSCS ``Horizon'' collaboration.  We acknowledge the
help of the XT4 support team at CSCS, most notably Timothy Stitt.
%%
The XT5 performance results have been obtained on a Cray XT5
supercomputer system, courtesy of Cray Inc.  We acknowledge the help of
Stefan Andersson (Cray Inc.).
%%
We thank Jonathan Hu and Raymond Tuminaro (Sandia National Laboratories)
for valuable information and help regarding the ML package.
%%
% Special thanks are addressed to reviewers for their valuable
% suggestions.

\bibliography{paper}
\bibliographystyle{elsarticle-num}

\end{document}
