\begin{thebibliography}{10}
%1 
\bibitem{bib:GKS}
{\sl The Graphical Kernel System (GKS)}.
ISO, Geneva, July 1985.
International Standard ISO 7942.
%2 
\bibitem{bib:autin}
B. Autin and Y. Marti.
{\sl Closed Orbit Correction of Alternating Gradient Machines
  using a small Number of Magnets}.
CERN/ISR-MA/73-17, CERN, 1973.
%3 
\bibitem{bib:barber} 
D.P.~Barber, K. Heinemann, H.~Mais and G.~Ripken,
  {\sl A Fokker--Planck Treatment of Stochastic Particle Motion within
    the Framework of a Fully Coupled 6-dimensional Formalism for
    Electron-Positron Storage Rings including Classical Spin Motion in
    Linear Approximation}, 
DESY report 91-146, 1991.
%4
\bibitem{bib:scandale} 
R.~Bartolini, A.~Bazzani, M.~Giovannozzi, W.~Scandale and E. Todesco, 
{\sl Tune evaluation in simulations and experiments}, 
CERN SL/95-84 (AP) (1995).
%5 
\bibitem{bib:bjorken}
J.~D.~Bjorken and S.~K.~Mtingwa.
{\sl Particle Accelerators} {\bf 13}, pg. 115.
%6 
\bibitem{bib:bollt} 
E.M.~Bollt and J.D.~Meiss, 
{\sl Targeting chaotic orbits to the Moon through recurrence}, 
Phys.~Lett. A{\bf 204}, 373 (1995).
%7 
\bibitem{B-bib:bramham}
P. Bramham and H. Henke.
private communication and LEP Note LEP-70/107, CERN.
%8 
\bibitem{bib:matrix}
Karl~L. Brown.
{\sl A First-and Second-Order Matrix Theory for the Design
  of Beam Transport Systems and Charged Particle Spectrometers}.
SLAC 75, Revision 3, SLAC, 1972.
%9 
\bibitem{bib:transport}
Karl~L. Brown, D.~C. Carey, Ch. Iselin, and F. Rothacker.
{\sl TRANSPORT --- A Computer Program for Designing Charged
  Particle Beam Transport Systems}.
CERN 73-16, revised as CERN 80-4, CERN, 1980.
%10 
\bibitem{bib:chao}
A. Chao.
{\sl Evaluation of beam distribution parameters in an electron
  storage ring}.
Journal of Applied Physics, 50:595--598, 1979.
%11
\bibitem{bib:touschek}
A.~W. Chao and M.~J. Lee.
{\sl SPEAR II Touschek lifetime}.
SPEAR-181, SLAC, October 1974.
%12 
\bibitem{bib:conte}
M.~Conte and M.~Martini.
{\sl Particle Accelerators} {\bf 17}, 1 (1985).
%13
\bibitem{bib:courant}
E.~D. Courant and H.~S. Snyder.
{\sl Theory of the alternating gradient synchrotron}.
Annals of Physics, 3:1--48, 1958.
%14 
\bibitem{bib:TFS}
Ph. Defert, Ph. Hofmann, and R. Keyser.
{\sl The Table File System, the C Interfaces}.
LAW Note 9, CERN, 1989.
%15 
\bibitem{bib:HARMON}
M. Donald and D. Schofield.
{\sl A User's Guide to the {\tt HARMON} Program}.
LEP Note 420, CERN, 1982.
%16 
\bibitem{bib:dragt}
A. Dragt.
{\sl Lectures on Nonlinear Orbit Dynamics, 1981 Summer School on High
  Energy Particle Accelerators, Fermi National Accelerator Laboratory, July
  1981}.
American Institute of Physics, 1982.
%17 
\bibitem{bib:edwards}
D.~A. Edwards and L.~C. Teng.
{\sl Parametrisation of linear coupled motion in periodic systems}.
IEEE Trans. on Nucl. Sc., 20:885, 1973.
%18  
\bibitem{bib:giovannozzi} 
M.~Giovannozzi, 
{\sl Analysis of the stability domain of planar symplectic maps using invariant manifolds}, 
CERN/PS 96-05~(PA) (1996).
%19 
\bibitem{bib:gxplot}
H. Grote.
{\sl GXPLOT User's Guide and Reference Manual}.
LEP TH Note 57, CERN, 1988.
%20 
\bibitem{bib:LEP}
LEP~Design Group.
{\sl Design Study of a 22 to 130 GeV e\({}^{+}\)e\({}^{-}\) Colliding Beam
  Machine (LEP)}.
CERN/ISR-LEP/79-33, CERN, 1979.
%21 
\bibitem{bib:BMPM}
M. Hanney, J.~M. Jowett, and E. Keil.
{\sl BEAMPARAM --- A program for computing beam dynamics and
  performance of $e^{+}e^{-}$ storage rings}.
CERN/LEP-TH/88-2, CERN, 1988.
%22 
\bibitem{bib:helm}
R.~H. Helm, M.~J. Lee, P.~L. Morton, and M. Sands.
{\sl Evaluation of synchrotron radiation integrals}.
IEEE Trans. Nucl. Sc., NS-20, 1973.
%23 
\bibitem{bib:MINUIT}
F. James.
{\sl MINUIT, A package of programs to minimise a function of n
  variables, compute the covariance matrix, and find the true errors}.
program library code D507, CERN, 1978.
%24 
\bibitem{bib:keil}
E. Keil.
{\sl Synchrotron radiation from a large electron-positron storage ring}.
CERN/ISR-LTD/76-23, CERN, 1976.
%25 
\bibitem{bib:knuth}
D.~E. Knuth.
{\sl The Art of Computer Programming}.
Volume~2, Addison-Wesley, second edition, 1981.
Semi-numerical Algorithms.
%26 
\bibitem{bib:laskar} 
J.~Laskar, C.~Froeschl\'{e} and A.~Celletti, 
{\sl The measure of chaos by the numerical analysis of the fundamental frequencies. Application to the standard mapping}, 
Physica D{\bf 56}, 253 (1992).
%27
\bibitem{bib:mais}
H.~Mais and G.~Ripken,
{\sl Theory of Coupled Synchro-Betatron Oscillations}.
DESY internal Report, DESY M-82-05, 1982.
%28
\bibitem{bib:meddahi} 
M.~Meddahi, 
{\sl Chromaticity correction for the $108^{\circ}/60^{\circ}$ lattice}, 
CERN SL/Note 96-19 (AP) (1996).
%29
\bibitem{bib:milutinovic}
J. Milutinovic and S. Ruggiero.
{\sl Comparison of Accelerator Codes for a RHIC Lattice}.
AD/AP/TN-9, BNL, 1988.
%30 
\bibitem{bib:montague}
B.~W. Montague.
{\sl Linear Optics for Improved Chromaticity Correction}.
LEP Note 165, CERN, 1979.
%31
\bibitem{bib:ripken}
G. Ripken,
{\sl Untersuchungen zur Strahlf\"uhrung und Stabilit\"at der Teilchenbewegung in Beschleunigern und Storage-Ringen unter strenger Ber\"ucksichtigung einer Kopplung der Betatronschwingungen}.
DESY internal Report R1-70/4, 1970.
%32 
\bibitem{bib:chamonix} 
F.~Ruggiero,
{\sl Dynamic Aperture for LEP~2 with various optics and tunes}, 
Proc. Sixth Workshop on LEP Performance, Chamonix, 1996, ed.~J.~Poole (CERN SL/96-05 (DI), 1996), pp.~132--136.
%33 
\bibitem{bib:teng}
L.~C. Teng.
{\sl Concerning n-Dimensional Coupled Motion}.
FN 229, FNAL, 1971.
%34 
\bibitem{bib:volkel}
U. V\"olkel.
{\sl Particle loss by Touschek effect in a storage ring}.
DESY 67-5, DESY, 1967.
%35 
\bibitem{bib:walker}
R.~P. Walker.
{\sl Calculation of the Touschek lifetime in electron storage rings}.
1987.
Also SERC Daresbury Laboratory preprint, DL/SCI/P542A.
%36 
\bibitem{bib:wilson}
P.~B. Wilson.
{\sl Proc. 8th Int. Conf. on High-Energy Accelerators}.
Stanford, 1974.
%37 
\bibitem{bib:wrulich}
A. Wrulich and H. Meyer.
{\sl Life time due to the beam-beam bremsstrahlung effect}.
PET-75-2, DESY, 1975.
%38
\bibitem{langdon}
Birdsall, Langdon.
\newblock {\em Plasma Physics via computer simulation}.
\newblock Page 340-341
%39
\bibitem{fubiani2006}
G.\ Fubaiani, J.\ Qiang, E.\ Esarey, W.\ P.\ Leemans, G.\ Dugan.
\newblock {\em Space charge modeling of dense electron beams with large energy spreads}.
\newblock Acclerators and Beams, 9, 064402 (2006) 
%40 
\bibitem{qiang2005}
J. Qiang, S. Lidia, R. D. Ryne, and C. Limborg-Deprey
\newblock {\em A Three-Dimensional Quasi-Static Model for High Brightness Beam Dynamics Simulation}
\newblock Lawrence Berkeley National Laboratory. Paper LBNL-59098. http://repositories.cdlib.org/lbnl/LBNL-59098
%41 
\bibitem{qiang2006-2}
J. Qiang, S. Lidia, R. D. Ryne, C. Limborg-Deprey, 
\newblock Phys. Rev. Special Topics - Accel. Beams 9, 044204, (2006).
%42 
\bibitem{JohoDist}
W. Joho
\newblock private discussions
%43
\bibitem{superfish}
J. H. Billen, L. M. Young
{\sl POISSON SUPERFISH}.
LA-UR-96-1834, Los Alamos National Laboratory, 2004
%44
\bibitem{enge}
J. E. Spencer,  H. A. Enge
\newblock {\em Split-Pole Magnetic Spectrograph for Precision Nuclear Spectroscopy}
\newblock Nucl. Instr. Meth. 49 (1967) 181-193
%45
\bibitem{SanzSerna}
J.M. Sanz-Sern and M.P. Calvo
\newblock {\em Numerical Hamiltonian Problems}
Chapman and Hall 1994
%46
\bibitem{forestall}
E. Forest. Phys. Lett. A 158, p99, 1991
%47
\bibitem{bib:arbenzetal2001}
P. Arbenz, R. Geus and S. Adam
{\sl Solving Maxwell eigenvalue problems for accelerating cavities,}
Physical Review Special Topics - Accelerators and Beams,
4, pp. 022001-1:022001-10, 2001.
%48
\bibitem{bib:arbenzetal2006}
P. Arbenz, M. Be\v{c}ka, R. Geus, U. Hetmaniuk and T. Mengotti
{\sl On a Parallel Multilevel Preconditioned Maxwell Eigensolver,}
PARCO, 32(2), pp. 157-165, 2006, doi: 10.1016/j.parco.2005.06.005
%49
\bibitem{flo:97}
K. Fl\"ottmann,
{\sl Note on the thermal emittance of electrons emitted by Cesium Telluride photo cathodes},
\newblock {\em TESLA FEL-Report}, 1997-01.
%50
\bibitem{clen:2000}
J. E. Clendenin, T. Kotseroglou, G. A. Mulhollan, D. T. Palmer, J. F. Schmerge,
{\sl Reduction of thermal emittance of RF guns}
\newblock {\em NIM A}, 455 (2000) 198-201.
%51
\bibitem{Adelmann:2009p543}
A Adelmann and P Arbenz and Y Ineichen, 
{\sl A Fast Parallel Poisson Solver on Irregular Domains Applied to Beam  Dynamic Simulations}
\newblock {\em 0907.4863v1 arXiv} (2009)
\bibitem{BC} Y. Feng and J. P. Verboncoeur,
{\sl Transition from Fowler-Nordheim field emission to space charge limited current density}
\newblock {\em Phys.Plasmas}, 13, 073105 (2006)
\bibitem{FN} R. H. Fowler and L. Nordheim, 
{\sl Electron Emission in Intense Electric Fields}
\newblock {\em Proc. R. Soc. London}, Ser. A 119, 173 (1928)
\bibitem{DE} J. H. Han, 
{\sl Dynamics of Electron Beam and Dark Current in Photocathode RF Guns}
\newblock PhD Thesis, Desy, 2005 Available Online on \url{http://www-library.desy.de/preparch/desy/thesis/desy-thesis-05-045.pdf}
\bibitem{WangHB2010} C. Wang and A. Adelmann and Y. Ineichen,
{\sl A Field Emission and Secondary Emission Model in \opal}
\newblock {\em Proc. of HB2010}, MOPD55 (2010) 
\bibitem{Furman-Pivi} M. A. Furman and M. T. F. Pivi,
{\sl Probabilistic model for the simulation of secondary electron emission}
\newblock {\em Phys. Rev. ST Accel. Beams}, 5, 124404 (2002) 
\bibitem{Vaughan} J. Rodney M. Vaughan,
{\sl A New Formula for Secondary Emission Yield}
\newblock {\em IEEE Trans. on Electron Devices}, 36:9, 1963 (1989)  
\bibitem{VaughanRv} J. Rodney M. Vaughan,
{\sl Secondary Emission Formulas}
\newblock {\em IEEE Trans. on Electron Devices}, 40:4, 830 (1993) 
\bibitem{FS} C. Vicente and M. Mattes and D. Wolk and H. L. Hartnagel and J. R. Mosig and D. Raboso,
{\sl FEST3D - A Simulation Tool for Multipactor Prediction}
\newblock {\em Proc. of MULCOPIM 2005}, (2005)  
\bibitem{nonstationary} S. Anza and C. Vicente and J. Gil and V. E. Boria and B. Gimeno and D. Raboso,
{\sl Nonstationary Statistical Theory for Multipactor}
\newblock {\em Phys. Plasmas}, 17, 062110 (2010)
\bibitem{paraview} A. Henderson,
{\sl ParaView Guide: A Parallel Visualization Application}
\newblock {\em Kitware Inc.}, (2007)  
\end{thebibliography}
